---
metadata-files: 
    - ../../_variables.yml
title: |
    <span class='fa-stack course-icon'>
        <i class='fas {{< var course.icon-bg >}} fa-stack-2x blue'></i>
        <i class='fas {{< var course.icon >}} fa-stack-1x fa-inverse' style='margin-top:-5pt'></i>
      </span><br>
    {{< var course.subtitle >}}
author: "{{< var instructor.name >}}"  
subtitle: Maps
date: December 4, 2023
date-format: long
from: markdown+emoji
bibliography: ../../data/lit/wipol_advanced.bib
csl: ../../data/lit/wipol.csl
execute:
  freeze: auto
  echo: true
  fig-width: 6
  fig-height: 5
format:
  revealjs:
    theme: ../econpol_slides.scss
    footer: '{{< var course.number >}} {{< var course.title >}} | {{< var course.semester >}}'
    logo: ../../data/img/wu-logo-simple.png
    header-includes: |
      <link href="https://cdnjs.cloudflare.com/ajax/libs/font-awesome/6.1.1/css/all.min.css" rel="stylesheet"/>
    slide-number: true
    show-slide-number: all
    transition: fade
    progress: true
    multiplex: false
    scrollable: false
    preview-links: false
    hide-inactive-cursor: true
    highlight-style: printing
    pause: true
---


## Projections {.medium}

Choose a [map projection]{.marker-hl}! For instance, [{giscoR}]{.secfont} by Eurostat provides maps in the following three projections. Check coordinates for axis limits at [https://epsg.io](https://epsg.io).

```{r projections}
#| echo: false
#| results: 'asis'
#| fig-width: 12
#| fig-align: center
library(tidyverse)
library(sf)
library(giscoR)
library(patchwork)

load("data/projections.RData")

map1 <- europe_4326 |> ggplot() + 
  geom_sf(fill = "gray95", linewidth = .1, color = "black") + 
  coord_sf(xlim = c(-26, 50), ylim = c(35, 72), expand = F) +
  labs(title = "EPSG:4326 World Geodetic System 1984 (WGS84)")
map2 <- europe_3035 |> ggplot() + 
  geom_sf(fill = "gray95", linewidth = .1, color = "black") + 
  coord_sf(xlim = c(2377294, 7453440), ylim = c(1313597, 5628510), expand = F) +
  labs(title = "EPSG:3035 ETRS89-extended / LAEA Europe")
map3 <- europe_3857 |> ggplot() + 
  geom_sf(fill = "gray95", linewidth = .1, color = "black") + 
  coord_sf(xlim = c(-3090000, 4500000), ylim = c(11500000,  4100000), expand = F) +
  labs(title = "EPSG:3857 Pseudo-Mercator")

map1 + map2 + map3 & theme_minimal(base_family = "Roboto Condensed") +
  theme(axis.text = element_text(size = 7),
        panel.grid = element_line(linewidth = 0.2, color = "gray80"),
        panel.background = element_rect(color = NA, fill = "aliceblue"),
        plot.title = element_text(size = 10))
```

## The {sf} package {.medium}

The [sf]{.secfont} (*simple features*) package deals with [spatial geometries]{.marker-hl}. These features have a geometry describing where on Earth the feature is located, and they have attributes, which describe other properties.

```{r sfgeom}
#| echo: false
#| results: 'asis'
#| fig-width: 11
#| fig-align: center
p <- rbind(c(3.2,4), c(3,4.6), c(3.8,4.4), c(3.5,3.8), c(3.4,3.6), c(3.9,4.5))
mp <- st_multipoint(p)
p1 <- mp |> ggplot() + geom_sf() + labs(title = "Multipoint")
s1 <- rbind(c(0,3),c(0,4),c(1,5),c(2,5))
s2 <- rbind(c(0.2,3), c(0.2,4), c(1,4.8), c(2,4.8))
s3 <- rbind(c(0,4.4), c(0.6,5))
mls <- st_multilinestring(list(s1,s2,s3))
p2 <- mls |> ggplot() + geom_sf() + labs(title = "Multilinestring")
mp1 <- rbind(c(0,0), c(1,0), c(3,2), c(2,4), c(1,4), c(0,0))
mp2 <- rbind(c(1,1), c(1,2), c(2,2), c(1,1))
mp3 <- rbind(c(3,0), c(4,0), c(4,1), c(3,1), c(3,0))
mp4 <- rbind(c(3.3,0.3), c(3.8,0.3), c(3.8,0.8), c(3.3,0.8), c(3.3,0.3))[5:1,]
mp5 <- rbind(c(3,3), c(4,2), c(4,3), c(3,3))
mpol <- st_multipolygon(list(list(mp1,mp2), list(mp3,mp4), list(mp5)))
p3 <- mpol |> ggplot() + geom_sf() + labs(title = "Multipolygon")
p1 + p2 + p3 & 
  theme_minimal(base_family = "Roboto Condensed", base_size = 12) + 
  theme(plot.title = element_text(hjust = 0.5),
        panel.background = element_rect(fill = "transparent"),
        plot.margin = margin(l = 1, r = 1, unit = "lines"))
```

## Plot a map {.smaller}

```{r plotmap}
#| layout-ncol: 2
library(giscoR)
europe <- gisco_get_countries(region = "Europe", resolution = "10", epsg = 3035)
europe |> select(CNTR_ID, NAME_ENGL, geometry) |> slice_head(n = 3)
europe |> ggplot() + 
  geom_sf(fill = "midnightblue", linewidth = .1, color = "white") + 
  coord_sf(xlim = c(2377294, 7453440), ylim = c(1313597, 5628510), expand = F)
```
