---
title: "The functional distribution of income"
author: "Matthias Schnetzer"
date: "2023-04-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(MetBrewer)
library(readxl)
load("07_income.RData")
```

### Wage share in Austria

The wage share in **Austria** has been decreasing since the 1970s. In the wake of the global financial crisis that erupted in 2008, the wage share increased due to a sharp decline in profits. 

```{r austria, echo = FALSE, fig.width=10, fig.height=5}
wsdata |> filter(country == "Austria") |>
  ggplot(aes(x=year, y=value)) +
  geom_line(linewidth = 1.5, color = "red") +
  labs(y = "Wage share in % of GDP at factor cost", x = NULL, caption = "Data: Ameco.") +
  theme_minimal() + 
  theme(legend.position = "none",
        plot.caption = element_text(size = 10, margin = margin(t = 1, unit = "lines")),
        axis.title = element_text(size = 12),
        axis.text = element_text(size = 10))
```

We can improve the figure by editing the scales of the axes. We add percentage symbols to the y-axis and we remove the white space on the x-axis.

```{r autwithscales, echo = FALSE, fig.width=10, fig.height=5}
wsdata |> filter(country == "Austria") |>
  ggplot(aes(x=year, y=value)) +
  geom_line(linewidth = 1.5, color = "red") +
  scale_x_date(breaks= seq(min(wsdata$year),max(wsdata$year), by = "10 years"), 
               date_labels = "%Y", expand = c(0,0)) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  labs(y = NULL, x = NULL, caption = "Data: Ameco.",
       title = "Wage share in Austria", subtitle = "Wage share in % of GDP at factor cost") +
  theme_minimal() + 
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        plot.title = element_text(size = 14),
        plot.caption = element_text(size = 10, margin = margin(t = 1, unit = "lines")),
        axis.title = element_text(size = 12),
        axis.text = element_text(size = 10))
```

### Wage shares in selected international countries

We can plot the wage shares for selected countries and emphasize the data for Austria with `gghighlight`. We can also improve the axis labels with scales.

```{r wageshares, echo=FALSE, fig.width=10, fig.height=5}
library(gghighlight)
library(ggrepel)
wsdata |>
  ggplot(aes(x = year, y = value, group = country, color = country)) +
  geom_line(linewidth = 1.4) +
  gghighlight(country == "Austria", 
              unhighlighted_params = list(color = "gray80", alpha = 0.5, linewidth = 0.9),
              line_label_type = "label_path", label_key = country, use_group_by = F,
              label_params = list(size = 3, hjust = 0.94)) +
  geom_text_repel(data = wsdata |> slice_max(year, n=1, by=country) |> filter(!country == "Austria"), 
                  aes(x = year, y = value, label = country), 
                  xlim = c(as.Date("2023-01-01"), NA),
                  size = 2.5, color = "gray70", hjust = 0, direction = "y",
                  min.segment.length = unit(2, "pt"), segment.size = 0.1) +
  scale_color_manual(name = NULL, values = met.brewer("Austria"),) +
  scale_x_date(breaks= seq(min(wsdata$year),max(wsdata$year), by = "10 years"), date_labels = "%Y",
               limits = c(as.Date("1960-01-01"), as.Date("2026-01-01"))) +
  scale_y_continuous(limits = c(55,80), labels = scales::percent_format(scale = 1)) +
  labs(y = NULL, x = NULL, caption = "Data: Ameco.",
       title = "Wage share in selected countries", subtitle = "Wage share in % of GDP at factor cost") +
  theme_minimal() + 
  theme(panel.grid.minor = element_blank(),
        plot.title = element_text(size = 14),
        plot.caption = element_text(size = 10, margin = margin(t = 1, unit = "lines")),
        axis.title = element_text(size = 12),
        axis.text = element_text(size = 10))
```

### Code for the figure

Here is the **clean code** for the figure on the functional distribution.

```{r wscode, echo=TRUE, eval=FALSE}
wsdata |>
  ggplot(aes(x = year, y = value, group = country, color = country)) +
  geom_line(linewidth = 1.4) +
  gghighlight(country == "Austria", 
              unhighlighted_params = list(color = "gray80", alpha = 0.5, linewidth = 0.9),
              line_label_type = "label_path", label_key = country, use_group_by = F,
              label_params = list(size = 3, hjust = 0.94)) +
  geom_text_repel(data = wsdata |> slice_max(year, n=1, by=country) |> filter(!country == "Austria"), 
                  aes(x = year, y = value, label = country), 
                  xlim = c(as.Date("2023-01-01"), NA),
                  size = 2.5, color = "gray70", hjust = 0, direction = "y",
                  min.segment.length = unit(2, "pt"), segment.size = 0.1) +
  scale_color_manual(name = NULL, values = met.brewer("Austria"),) +
  scale_x_date(breaks= seq(min(wsdata$year),max(wsdata$year), by = "10 years"), date_labels = "%Y",
               limits = c(as.Date("1960-01-01"), as.Date("2026-01-01"))) +
  scale_y_continuous(limits = c(55,80), labels = scales::percent_format(scale = 1)) +
  labs(y = NULL, x = NULL, caption = "Data: Ameco.",
       title = "Wage share in selected countries", subtitle = "Wage share in % of GDP at factor cost") +
  theme_minimal() + 
  theme(panel.grid.minor = element_blank(),
        plot.title = element_text(size = 14),
        plot.caption = element_text(size = 10, margin = margin(t = 1, unit = "lines")),
        axis.title = element_text(size = 12),
        axis.text = element_text(size = 10))
```